### 2017-04-13 Papers/coverletter.tex

\documentclass[12pt]{article}
\usepackage[a4paper,margin=1cm,footskip=.5cm]{geometry}
\usepackage[]{authblk}
\usepackage{graphicx}
\usepackage{hanging}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{multicol}
\usepackage{color}
\usepackage{lipsum}
\usepackage[breaklinks]{hyperref}
\usepackage[all]{hypcap}
\definecolor{darkblue}{rgb}{0.0,0.0,0.5}
%\onehalfspace
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.75in}
\setlength{\tabcolsep}{0in}
\setlength{\textheight}{9in} % This changed from 9in
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{0.15in}
\setlength{\topskip}{0mm}
\setlength{\voffset}{-0.5in}
\setlength{\parindent}{1cm}
%\setlength{\headheight}{99pt}
\setlength{\headheight}{0.5in}
\parskip = 1mm
\pagestyle{plain}

\definecolor{citescol}{RGB}{73,0,165}
\definecolor{urlscol}{RGB}{3,0,255}
\definecolor{linkscol}{RGB}{187,24,0}
\definecolor{isured}{RGB}{204,0,51}
\definecolor{isukhaki}{RGB}{76,69,43}

\hypersetup{colorlinks=true,linkcolor=linkscol,citecolor=citescol,urlcolor=urlscol}

\lhead{{\textcolor{isured}{\LARGE{\textsc{Iowa State University}}}}\\{\textcolor{isukhaki}{OF SCIENCE AND TECHNOLOGY}}}
\rhead{\scriptsize{\includegraphics[width=0.85in]{./assets/EEOB-eps-converted-to}\\{\textcolor{isured}{Ecology, Evolution, and Organismal Biology}}}}

\begin{document}

Dear Surayya,

Thank you very much for your prompt handling of my manuscript ``Editor's Note on `Putting Fossils in Trees' Special Issue". In this revision, I have addressed the comments submitted by the reviewer.

Thank you for your time, and I look forward to hearing from you.

\vspace{6mm}
\noindent Sincerely,\\
\includegraphics[width=2in]{./assets/Sig2}\\
\noindent Dr. April M. Wright\\
Postdoctoral Researcher\\
Dept. Ecology, Evolution, \& Organismal Biology\\
Iowa State University\\
251 Bessey Hall\\
Ames, IA 50011\\
\href{mailto:wright.aprilm@gmail.com}{{wright.aprilm@gmail.com}}\\
(512) 940-5761

\end{document}